---
title: Loading packages
jupyter:
  kernelspec:
    display_name: R
    name: ir
    language: R
---

```{r}
library(tidyverse)
install.packages("GGally")
library(GGally)
library(readr)
library(stringr)
install.packages("modeest")
library(modeest)
install.packages("gridExtra")
library(grid)
library(gridExtra)
install.packages("ggthemes")
library(ggthemes)
install.packages("shiny")
library(shiny)
install.packages("ggcorrplot")
library(ggcorrplot)
install.packages("car")
library(car)
install.packages("ggfortify")
library(ggfortify)
install.packages("randomForest")
library(randomForest)
library(scales)
install.packages("fastDummies")
library(fastDummies)
install.packages("caret")
library(caret)
install.packages("remotes")
remotes::install_github("ltorgo/DMwR2", ref = "develop")
install.packages("DMwR2")
library(DMwR2)
install.packages("mice")
library(mice)
library(reshape2)
install.packages("readr")
library(readr)
```

# Loading data

```{r}
Car_df = read_csv("https://raw.githubusercontent.com/NitoBoritto/R_New_York_Car_Project/main/New_York_cars.csv")
Car_rates = read_csv("https://raw.githubusercontent.com/NitoBoritto/R_New_York_Car_Project/main/Car_Rates.csv")
```

```{r}
Car_df %>% head(3)
```

```{r}
Car_df %>% str()
```

```{r}
Car_rates %>% head()
```

```{r}
Car_rates %>% str()
```

# Preprocessing

## Checking Duplicates

```{r}
Dups_Car = Car_df %>%
  duplicated() %>%
  sum()
sprintf("Duplicates in Car dataset before: %d", Dups_Car)

Dups_rate = Car_rates %>%
  duplicated() %>%
  sum()
sprintf("Duplicates in Rate dataset: %d", Dups_rate)
```

```{r}
# Deleting empty rows
Car_df = Car_df %>% distinct()

Dups_Car_00 = Car_df %>%
  duplicated() %>%
  sum()
sprintf("Duplicates in Car dataset after: %d", Dups_Car_00)
```

## `name` Column

No nulls

```{r}
name_nulls = Car_df$name %>%
  is.na() %>%
  sum()

sprintf("Nulls: %d", name_nulls)
```

```{r}
Car_df$name %>% unique()
```

```{r}
Car_rates$Car_name %>% unique()
```

```{r}
Car_df = Car_df %>%
  mutate(
    # Step 1: Extract the primary model name from the Model column
    Primary_Model = word(Model, 1), # Extracts the first word as the primary model
    # Step 2 & 3: Combine Year, brand, and primary model name, then trim whitespace
    Car_df_join_name = paste(as.character(Year), brand, Primary_Model, sep = " ") %>%
                       str_trim(),
    # Step 4: Append a single period at the end
    Car_df_join_name = paste0(Car_df_join_name, ".")
  )

print("First few rows of Car_df with the new Car_df_join_name:")
Car_df %>% select(name, Model, Car_df_join_name) %>% head()
```

```{r}
Car_rates = Car_rates %>%
  mutate(
    Car_name_cleaned = Car_name %>%
                       str_replace_all(" +", " ") %>% # Replace multiple spaces with single space
                       str_trim() %>% # Trim leading/trailing whitespace
                       str_replace("\\.$|^$", "") %>% # Remove any existing period at the end
                       paste0(".") # Append a single period at the end
  )

# Display the original Car_name and the new Car_name_cleaned for verification
print("First few rows of Car_rates with the original and cleaned Car_name:")
Car_rates %>%
  select(Car_name, Car_name_cleaned) %>%
  head()
```

```{r}
Car_df_merged = Car_df %>%
  left_join(Car_rates, by = c("Car_df_join_name" = "Car_name_cleaned"), suffix = c("_car", "_rate"))

Car_df_merged = Car_df_merged %>%
  select(
    -Year_rate, -Brand, -Model_rate, -Car_name, # Columns from Car_rates to remove
    -name, # Intermediate columns from Car_df (Primary_Model not suffixed as no conflict)
    -`Interior color`, -`Exterior color`, -Convenience, -Entertainment, # Other Car_df columns to remove (not suffixed)
    -Exterior, -Safety, -Seating
  ) %>%
  rename(
    Year = Year_car, # Rename Year from Car_df (suffixed as Year_car) back to Year
    Model = Model_car, # Rename Model from Car_df (suffixed as Model_car) back to Model
    name = Car_df_join_name
  )

Car_df_merged %>% head()

print("Dimensions of the merged DataFrame:")
Car_df_merged %>% dim()
```

## `brand` Column

No nulls

```{r}
brand_nulls = Car_df_merged$brand %>%
  is.na %>%
  sum()

sprintf("Nulls: %d", brand_nulls)
```

```{r}
brand_unique = Car_df_merged$brand %>%
  unique() %>%
  length()
cat("There are", brand_unique, "brands of cars in NY")
```

```{r}
Column_names = Car_df_merged %>% colnames()

Car_df_merged %>%
  semi_join(Car_rates, by = c("brand" = "Brand")) %>%
  count()
```

Checking data inconsistency between 2 tables

```{r}
print('Brands in Car_df not matching Car_rates (anti_join):')
Car_df_merged %>%
  anti_join(Car_rates, by = c("brand" = "Brand")) %>%
  pull(brand) %>%
  unique()
```

```{r}
print('Brands in Car_rates not matching Car_df (anti_join):')
Car_rates %>%
  anti_join(Car_df_merged, by = c("Brand" = "brand")) %>%
  pull(Brand) %>%
  unique()
```

```{r}
# Unifying brand names
Car_df_merged = Car_df_merged %>%
  mutate(brand = str_replace_all(brand,
                  c("Bmw" = "BMW",
                  "Gmc" = "GMC",
                  "Infiniti" = "INFINITI",
                  "Land_Rover" = "Land-Rover",
                  "Mercedes_Benz" = "Mercedes-Benz",
                  "Ram" = "RAM")))


Car_rates = Car_rates %>%
  mutate(Brand = str_replace(Brand, "Land", "Land-Rover"))
```

```{r}
n_car = Car_df_merged %>%
  semi_join(Car_rates, by = c("brand" = "Brand")) %>%
  nrow()

sprintf('Rows in Car_df matching Car_rates by brand: %d', n_car)
```

```{r}
n_rates = Car_rates %>%
  semi_join(Car_df_merged, by = c("Brand" = "brand")) %>%
  nrow()

sprintf('Rows in Car_rates matching Car_df by brand: %d', n_rates)
```

## `new&used` Column

No nulls

```{r}
new_used_nulls =Car_df_merged$`new&used` %>%
  is.na() %>%
  sum()

sprintf("Nulls: %d", new_used_nulls)
```

A ton of inconsistent data

```{r}
Car_df_merged$`new&used` %>% table()
```

```{r}
# Setting up mode
mode_val = mfv(Car_df_merged$`new&used`)

# Imputation
Car_df_merged = Car_df_merged %>%
  mutate(`new&used` = ifelse(!`new&used` %in% c("Used", "New"), mode_val, `new&used`))

new_used_vals = Car_df_merged$`new&used` %>% unique()
cat("Column after impuration: [", paste(new_used_vals, collapse = ", "), "]")
```

## `money` column

No nulls

```{r}
money_nulls = Car_df_merged$money %>%
  is.na() %>%
  sum()

sprintf("Nulls: %d", money_nulls)
```

```{r}
options(repr.plot.width = 18, repr.plot.height = 10)

theme_nito = theme(plot.title = element_text(hjust = 0.5, color = "darkgreen"),
    text = element_text(family = "garamond", size = 14),
	  rect = element_blank(),
	  panel.grid = element_blank(),
	  axis.line = element_line(color = "black"))


money_charts = Car_df_merged %>%
  ggplot(aes(money)) + theme_nito +
  theme(axis.line.x = element_blank())

raw_money = money_charts +
  geom_histogram(bins = 30, fill = "blue", color = "black") +
  labs(title = "Raw Money Distribution",
     y = "Frequency",
     x = "Money")


logged_money = money_charts +
  geom_histogram(bins = 30, fill = "blue", color = "black") +
  scale_x_log10() +
  labs(title = "Logged Money Distribution",
     y = "Frequency",
     x = "Money")

grid.arrange(raw_money, logged_money, ncol = 2,
             top = textGrob("Spread of money vs logged money",
                            gp = gpar(fontsize = 16, col = "blue", fontface = "bold")))
```

## `Drivetrain` column

552 Nulls

```{r}

drive_nulls = Car_df_merged$Drivetrain %>%
  is.na() %>%
  sum()

sprintf("Nulls: %d", drive_nulls)
```

```{r}
# Identifying unqiue vals
drivetrain_vals = Car_df_merged$Drivetrain %>% unique()
print(Car_df$Drivetrain %>% unique())

Car_df_merged$Drivetrain %>% table()
```

```{r}
mode_drive_train = mfv(Car_df_merged$Drivetrain)

# Filtering for not null
drivetrain_notnull = Car_df_merged$Drivetrain[] %>%
  na.omit() %>%
  unique()

# Mode imputation
Car_df_merged = Car_df_merged %>%
  mutate(Drivetrain = ifelse(!Drivetrain %in% drivetrain_notnull, mode_drive_train, Drivetrain))

Car_df_merged$Drivetrain %>% table()
```

## `Fuel type` column

```{r}
fuel_type_nulls = Car_df_merged$`Fuel type` %>%
  is.na() %>%
  sum()

sprintf("Nulls: %d", fuel_type_nulls)
```

```{r}
print("Unique Fuel Types before encoding:")
original_fuel_types = Car_df_merged$`Fuel type` %>%
  unique()

original_fuel_types %>% print()

Car_df_merged$`Fuel type` %>% table()
```

```{r}
Car_df_merged = Car_df_merged %>%
  mutate(Fuel_type_encoded = case_when(
    `Fuel type` == "Diesel" ~ 1,
    `Fuel type` %in% c("Gasoline", "Gasolin") ~ 2,
    `Fuel type` == "E85 Flex" ~ 3,
    `Fuel type` == "Hybrid" ~ 4,
    !is.na(`Fuel type`) ~ 0,
    is.na(`Fuel type`) ~ NA_real_
  ))

print("Car_df_merged head with Fuel_type_encoded:")
Car_df_merged %>% select(`Fuel type`, Fuel_type_encoded) %>% head()
```

```{r}
fuel_encode_unique = Car_df_merged$Fuel_type_encoded %>%
  unique() %>%
  sort()
cat("Unique values in Fuel_type_encoded after initial encoding: [",
    paste(fuel_encode_unique, collapse = ", "), "]")


na_count_encoded = Car_df_merged$Fuel_type_encoded %>%
  is.na() %>%
  sum()
sprintf("Number of NA values in Fuel_type_encoded: %d", na_count_encoded)
```

```{r}
# Prepare data for kNN imputation:
# - Convert Fuel_type_encoded to factor for categorical imputation
# - Convert Drivetrain to factor as it's a categorical predictor
Car_df_for_imputation = Car_df_merged %>%
  mutate(
    Drivetrain = as.factor(Drivetrain),
    Fuel_type_encoded = as.factor(as.character(Fuel_type_encoded)) # Convert to character then factor to preserve integer levels
  ) %>%
  select(money, Mileage, Year, Drivetrain, Fuel_type_encoded)

set.seed(42)

# Perform kNN imputation. Only Fuel_type_encoded has NAs among selected columns.
imputed_data_full = knnImputation(Car_df_for_imputation, k = 5, meth = "median")

# Update the original Car_df_merged with the imputed values
Car_df_merged$Fuel_type_encoded = as.numeric(as.character(imputed_data_full$Fuel_type_encoded))


na_count_after_imputation = Car_df_merged$Fuel_type_encoded %>%
  is.na() %>%
  sum()
sprintf("Number of NA values in Fuel_type_encoded after KNN imputation: %d", na_count_after_imputation)

unique_fuel_after_imputation = Car_df_merged$Fuel_type_encoded %>%
  unique() %>%
  sort()
cat("Unique values in Fuel_type_encoded after KNN imputation: [",
    paste(unique_fuel_after_imputation, collapse = ", "), "]")
```

```{r}
# Decode the numerical values back to descriptive string labels
Car_df_merged = Car_df_merged %>%
  mutate(Fuel_type_decoded = case_when(
    Fuel_type_encoded == 0 ~ "Other",
    Fuel_type_encoded == 1 ~ "Diesel",
    Fuel_type_encoded == 2 ~ "Gasoline",
    Fuel_type_encoded == 3 ~ "E85 Flex",
    Fuel_type_encoded == 4 ~ "Hybrid",
  ))

print("Car_df_merged head with Fuel_type_decoded:")
Car_df_merged %>% select(`Fuel type`, Fuel_type_encoded, Fuel_type_decoded) %>% head()
```

```{r}
print("Frequency table for Fuel_type_decoded:")
Car_df_merged$Fuel_type_decoded %>% table()
```

```{r}
Car_df_merged = Car_df_merged %>%
  select(-Fuel_type_encoded, -`Fuel type`) %>%
  rename(Fuel_type = Fuel_type_decoded)

# Verify that the column has been removed
print("Columns in Car_df_merged after removing Fuel_type_encoded:")
Car_df_merged %>%
  colnames() %>%
  print()
```

## `Transmission` column

755 nulls

```{r}
trans_nulls = Car_df_merged$Transmission %>%
  is.na() %>%
  sum()

sprintf("Nulls: %d", trans_nulls)
```

```{r}
print("Unique Transmission Types before encoding:")
Car_df_merged$Transmission %>%
  unique()
```

```{r}
Car_df_merged = Car_df_merged %>%
  mutate(Transmission_encoded_temp = case_when(
    str_detect(str_to_lower(Transmission), "manual") ~ 1,
    str_detect(str_to_lower(Transmission), "automatic|auto|autostick|tiptronic|pdk|dct|geartronic|sportmatic|shiftronic") ~ 2,
    str_detect(str_to_lower(Transmission), "variable|cvt|xtronic|lineartronic|ivt|ecvt") ~ 3,
    !is.na(Transmission) ~ 0, # Map other non-NA strings to 0
    is.na(Transmission) ~ NA_real_
  ))

print("Car_df_merged head with Transmission_encoded_temp:")
Car_df_merged %>%
  select(Transmission, Transmission_encoded_temp) %>%
  head()
```

```{r}
Car_df_merged = Car_df_merged %>%
  mutate(Transmission_for_imputation = ifelse(
    Transmission_encoded_temp == 0 | is.na(Transmission_encoded_temp),
    NA_real_,
    Transmission_encoded_temp
  ))

Car_df_merged %>%
  select(Transmission, Transmission_encoded_temp, Transmission_for_imputation) %>%
  head()

print("Unique values in Transmission_for_imputation:")
Car_df_merged$Transmission_for_imputation %>%
  unique() %>%
  sort()

na_count_imputation = Car_df_merged$Transmission_for_imputation %>%
  is.na() %>%
  sum()
sprintf("Number of NA values in Transmission_for_imputation: %d", na_count_imputation)
```

```{r}
Car_df_for_transmission_imputation = Car_df_merged %>%
  mutate(
    Drivetrain = as.factor(Drivetrain),
    Fuel_type = as.factor(Fuel_type),
    Transmission_for_imputation = as.factor(as.character(Transmission_for_imputation))
  ) %>%
  select(money, Mileage, Year, Drivetrain, Fuel_type, Transmission_for_imputation)

set.seed(42)

imputed_transmission_data = knnImputation(Car_df_for_transmission_imputation, k = 5, meth = "median")

Car_df_merged$Transmission_imputed = as.numeric(as.character(imputed_transmission_data$Transmission_for_imputation))

na_count_after_imputation = Car_df_merged$Transmission_imputed %>%
  is.na() %>%
  sum()
sprintf("Number of NA values in Transmission_imputed after KNN imputation: %d", na_count_after_imputation)

unique_transmission_after_imputation = Car_df_merged$Transmission_imputed %>%
  unique() %>%
  sort()
cat("Unique values in Transmission_imputed after KNN imputation: [",
    paste(unique_transmission_after_imputation, collapse = ", "), "]")
```

```{r}
Car_df_merged = Car_df_merged %>%
  mutate(Transmission_decoded = case_when(
    Transmission_imputed == 1 ~ "Manual",
    Transmission_imputed == 2 ~ "Automatic",
    Transmission_imputed == 3 ~ "Variable",
  ))

print("Car_df_merged head with Transmission_imputed and Transmission_decoded:")
Car_df_merged %>%
  select(Transmission_imputed, Transmission_decoded) %>%
  head()
```

```{r}
Car_df_merged = Car_df_merged %>%
  select(-Transmission, -Transmission_encoded_temp, -Transmission_for_imputation, -Transmission_imputed) %>%
  rename(Transmission = Transmission_decoded)

print("Columns in Car_df_merged after updating Transmission and removing intermediate columns:")
Car_df_merged %>%
  colnames() %>%
  print()

print("Frequency table for the updated Transmission column:")
Car_df_merged$Transmission %>%
  table()
```

## `Engine` column

330 nulls

```{r}
engine_nulls = Car_df_merged$Engine %>%
  is.na() %>%
  sum()

sprintf("Nulls: %d", engine_nulls)
```

```{r}
print("Unique values in the 'Engine' column:")
Car_df$Engine %>% unique() %>% print()
```

```{r}
Car_df_merged = Car_df_merged %>%
  filter(!is.na(Engine))

engine_nulls_after = Car_df_merged$Engine %>%
  is.na() %>%
  sum()

sprintf("Nulls in Engine column after removal: %d", engine_nulls_after)
```

```{r}
Car_df_merged = Car_df_merged %>%
  mutate(
    # Refine regex to capture Engine Size (e.g., 1.5L, 3.5L, 2.0 Liter)
    Engine_Size = as.numeric(str_extract(`Engine`, "\\d+\\.\\d+(?=\\s*(L|Liter))|\\d+(?=\\s*(L|Liter))")),
    # Extract Cylinder Configuration (e.g., I4, V6, H6, 4-Cylinder)
    Cylinder_Configuration = case_when(
      str_detect(`Engine`, "(?i)I-?3|3.?Cylinder") ~ "I3",
      str_detect(`Engine`, "(?i)I-?4|4.?Cylinder") ~ "I4",
      str_detect(`Engine`, "(?i)I-?5|5.?Cylinder") ~ "I5",
      str_detect(`Engine`, "(?i)I-?6|6.?Cylinder") ~ "I6",
      str_detect(`Engine`, "(?i)V-?6") ~ "V6",
      str_detect(`Engine`, "(?i)V-?8") ~ "V8",
      str_detect(`Engine`, "(?i)V-?10") ~ "V10",
      str_detect(`Engine`, "(?i)V-?12") ~ "V12",
      str_detect(`Engine`, "(?i)H-?4|Flat.?4") ~ "H4",
      str_detect(`Engine`, "(?i)H-?6|Flat.?6") ~ "H6",
      TRUE ~ "Other/Unknown"
    )
  )

# Count missing values for new columns
print("Missing values in new columns:")
Car_df_merged %>%
  summarise(
    NA_Engine_Size = sum(is.na(Engine_Size)),
    NA_Cylinder_Configuration = sum(is.na(Cylinder_Configuration))
  ) %>% print()
```

```{r}
Car_df_merged = Car_df_merged %>%
  mutate(
    Has_Turbo = ifelse(str_detect(str_to_lower(Engine), "turbo|twin turbo"), 1, 0),
    Has_GDI = ifelse(str_detect(str_to_lower(Engine), "gdi|direct injection|pdi"), 1, 0),
    Has_DOHC = ifelse(str_detect(str_to_lower(Engine), "dohc"), 1, 0),
    Has_Supercharged = ifelse(str_detect(str_to_lower(Engine), "supercharged"), 1, 0)
  )

# Display the first few rows with the new columns
print("Car_df_merged head with extracted engine technology indicators:")
Car_df_merged %>%
  select(Engine, Engine_Size, Cylinder_Configuration, Has_Turbo, Has_GDI, Has_DOHC, Has_Supercharged) %>%
  head()

# Count missing values for new columns
print("Missing values in new technology columns:")
Car_df_merged %>%
  summarise(
    NA_Has_Turbo = sum(is.na(Has_Turbo)),
    NA_Has_GDI = sum(is.na(Has_GDI)),
    NA_Has_DOHC = sum(is.na(Has_DOHC)),
    NA_Has_Supercharged = sum(is.na(Has_Supercharged))
  ) %>%
  print()
```

```{r}
Car_df_for_engine_imputation = Car_df_merged %>%
  mutate(
    Drivetrain = as.factor(Drivetrain),
    Fuel_type = as.factor(Fuel_type),
    Transmission = as.factor(Transmission),
    Cylinder_Configuration = as.factor(Cylinder_Configuration)
  ) %>%
  select(money, Mileage, Year, Engine_Size, Drivetrain, Fuel_type, Transmission, Cylinder_Configuration)

print("Structure of Car_df_for_engine_imputation:")
Car_df_for_engine_imputation %>% str()

n_nulls_engine_size = Car_df_for_engine_imputation %>%
  filter(is.na(Engine_Size)) %>%
  count() %>%
  pull(n)
sprintf("Row count of NA in Engine_Size: %d", n_nulls_engine_size)
```

```{r}
set.seed(42)
imputed_engine_data = knnImputation(Car_df_for_engine_imputation, k = 5, meth = "median")

# Update the original Car_df_merged with the imputed values
Car_df_merged$Engine_Size = imputed_engine_data$Engine_Size

na_count_after_imputation = Car_df_merged$Engine_Size %>% is.na() %>% sum()
sprintf("Number of NA values in Engine_Size after KNN imputation: %d", na_count_after_imputation)

print("Car_df_merged head with imputed Engine_Size:")
Car_df_merged %>% select(Engine_Size, Cylinder_Configuration) %>% head()
```

```{r}
Car_df_merged = Car_df_merged %>% select(-Engine)

print("Columns in Car_df_merged after removing the 'Engine' column:")
Car_df_merged %>%
  colnames() %>%
  print()
```

## `Accidents or damage` column

52164 nulls

```{r}
accidents_null = Car_df_merged$`Accidents or damage` %>%
  is.na() %>%
  sum()

sprintf("Nulls: %d", accidents_null)
```

```{r}
Car_df_merged$`Accidents or damage` %>% table()
```

```{r}
# Encoding accidents as 1, the rest 0
Car_df_merged = Car_df_merged %>%
  mutate(`Accidents or damage` = as.integer(ifelse(`Accidents or damage` == "None Reported" | is.na(`Accidents or damage`) , 0, 1)))

print("After encoding and handling missing values:")
Car_df_merged$`Accidents or damage` %>% table()
```

## `Clean Title`

157340 nulls god damn

```{r}
clean_null = Car_df_merged$`Clean title` %>%
  is.na() %>%
  sum()

sprintf("Nulls: %d", clean_null)
```

```{r}
Car_df$`Clean title` %>% table()
```

```{r}
# Encoding clean title as 1, the rest 0
Car_df_merged = Car_df_merged %>%
  mutate(`Clean title` = as.integer(ifelse(`Clean title` == "No" | is.na(`Clean title`) , 0, 1)))

print("After encoding and handling missing values:")
Car_df_merged$`Clean title` %>% table()
```

## `1-owner vechile` column

52799 nulls

```{r}
owner_null = Car_df_merged$`1-owner vehicle` %>%
  is.na() %>%
  sum()

sprintf("Nulls: %d", owner_null)
```

```{r}
Car_df_merged$`1-owner vehicle` %>% table()
```

```{r}
# Encoding 1-owner as 1, the rest 0
Car_df_merged = Car_df_merged %>%
  mutate(`1-owner vehicle` = as.integer(ifelse(`1-owner vehicle` == "No" | is.na(`1-owner vehicle`) , 0, 1)))

print("After encoding and handling missing values:")
Car_df_merged$`1-owner vehicle` %>% table()
```

## `Personal use only` column

52575 nulls

```{r}
personal_null = Car_df_merged$`Personal use only` %>%
  is.na() %>%
  sum()

sprintf("Nulls: %d", personal_null)
```

```{r}
Car_df_merged$`Personal use only` %>% table()
```

```{r}
# Encoding personal use as 1, the rest 0
Car_df_merged = Car_df_merged %>%
  mutate(`Personal use only` = as.integer(ifelse(`Personal use only` == "No" | is.na(`Personal use only`) , 0, 1)))

print("After encoding and handling missing values:")
Car_df_merged$`Personal use only` %>% table()
```

## `Model` column

The focus for this column is to match it with rates dataset to join them later

```{r}
model_nulls = Car_df_merged$Model %>%
  is.na() %>%
  sum()

sprintf("Nulls: %d", model_nulls)
```

## `Mileage` column

```{r}

mileage_nulls = Car_df_merged$Mileage %>%
  is.na() %>%
  sum()

sprintf("Nulls: %d", mileage_nulls)
```

```{r}
# Calculate the mean of the non-missing Mileage values
mean_mileage = Car_df_merged$Mileage %>%
  na.omit() %>% # Remove NAs to calculate mean
  mean()

sprintf("Mean Mileage: %.2f", mean_mileage)

# Impute NA values in 'Mileage' with the calculated mean
Car_df_merged = Car_df_merged %>%
  mutate(Mileage = ifelse(is.na(Mileage), mean_mileage, Mileage))

# Verify that there are no more nulls in the 'Mileage' column
mileage_nulls_after_imputation = Car_df_merged$Mileage %>%
  is.na() %>%
  sum()
sprintf("Nulls in Mileage after imputation: %d", mileage_nulls_after_imputation)
```

## `MPG` column

```{r}
Car_df_merged$MPG %>%
  is.na() %>%
  sum()
```

```{r}
Car_df_merged$MPG %>%
  unique()
```

```{r}
Car_df_merged = Car_df_merged %>%
  mutate(MPG_cleaned = case_when(
    str_detect(MPG, "\\d+[-–-]\\d+") ~ (
      as.numeric(str_extract(MPG, "\\d+(?=[-–-])")) +
      as.numeric(str_extract(MPG, "(?<=[-–-])\\d+"))
    ) / 2,
    # Handles single numerical values (e.g., '30')
    str_detect(MPG, "^\\d+$") ~ as.numeric(MPG),
    TRUE ~ NA_real_
  ))

# Handle potential outliers that are parsed as very large numbers
# A common upper limit for MPG is around 100-150 for non-electric cars, marking anything above this as NA
Car_df_merged = Car_df_merged %>%
  mutate(MPG_cleaned = ifelse(MPG_cleaned > 150, NA_real_, MPG_cleaned))

print("Summary of MPG_cleaned after initial extraction and outlier handling:")
Car_df_merged$MPG_cleaned %>% summary()

na_count_mpg_cleaned = Car_df_merged$MPG_cleaned %>% is.na() %>% sum()
sprintf("Number of NA values in MPG_cleaned after initial cleaning: %d", na_count_mpg_cleaned)
```

```{r}
median_mpg = Car_df_merged$MPG_cleaned %>%
  na.omit() %>%
  median()

sprintf("Median MPG for imputation: %.2f", median_mpg)

Car_df_merged = Car_df_merged %>%
  mutate(MPG_cleaned = ifelse(is.na(MPG_cleaned), median_mpg, MPG_cleaned))

na_count_after_imputation = Car_df_merged$MPG_cleaned %>% is.na() %>% sum()
sprintf("Number of NA values in MPG_cleaned after imputation: %d", na_count_after_imputation)

print("Summary of MPG_cleaned after imputation:")
Car_df_merged$MPG_cleaned %>% summary()
```

```{r}
Car_df_merged = Car_df_merged %>%
  select(-MPG) %>%
  rename(MPG = MPG_cleaned)

print("First few rows of Car_df_merged with finalized MPG column:")
Car_df_merged %>% select(MPG) %>% head()

print("Columns in Car_df_merged after finalizing MPG:")
Car_df_merged %>% colnames()
```

## Saving the dataset before heavy imputation

```{r}
write_csv(Car_df_merged, "car_df_merged.csv")
```

## Imputation

### `Num_reviews`, `General_rate`, `Comfort`, `Interior design` `Performance`, `Value for the money`, `Exterior styling`, `Reliability` columns

```{r}
Car_df_merged = read_csv("https://raw.githubusercontent.com/NitoBoritto/R_New_York_Car_Project/main/car_df_merged.csv")
```

```{r}
Car_df_merged = Car_df_merged %>%
  mutate(
    Num_of_reviews = ifelse(Num_of_reviews < 1, NA, Num_of_reviews),
    General_rate = ifelse(General_rate < 0 | General_rate > 5, NA, General_rate),
    Comfort = ifelse(Comfort < 0 | Comfort > 5, NA, Comfort),
    `Interior design` = ifelse(`Interior design` < 0 | `Interior design` > 5, NA, `Interior design`),
    Performance = ifelse(Performance < 0 | Performance > 5, NA, Performance),
    `Value for the money` = ifelse(`Value for the money` < 0 | `Value for the money` > 5, NA, `Value for the money`),
    `Exterior styling` = ifelse(`Exterior styling` < 0 | `Exterior styling` > 5, NA, `Exterior styling`),
    Reliability = ifelse(Reliability < 0 | Reliability > 5, NA, Reliability)
  )

rating_cols = c("Num_of_reviews", "General_rate", "Comfort", "Interior design", "Performance", "Value for the money", "Exterior styling", "Reliability")

cat("\n--- After re-standardization ---\n")
for (col_name in rating_cols) {
  cat(sprintf("\n--- Column: %s ---\n", col_name))

  # Summary statistics
  summary_stats = Car_df_merged %>% pull(!!sym(col_name)) %>% summary()
  print(summary_stats)

  # NA count
  na_count = Car_df_merged %>% pull(!!sym(col_name)) %>% is.na() %>% sum()
  cat(sprintf("Number of NA values: %d\n", na_count))
}
```

```{r}
predictor_cols = c("money", "Mileage", "Year", "Engine_Size", "Drivetrain",
                   "Fuel_type", "Transmission", "Cylinder_Configuration",
                   "Has_Turbo", "Has_GDI", "Has_DOHC", "Has_Supercharged")

cat("\n--- Imputing Num_of_reviews ---\n")

# Create a focused temporary DataFrame for Num_of_reviews imputation
temp_df_num_reviews = Car_df_merged %>%
  mutate(
    Drivetrain = as.factor(Drivetrain),
    Fuel_type = as.factor(Fuel_type),
    Transmission = as.factor(Transmission),
    Cylinder_Configuration = as.factor(Cylinder_Configuration)
  ) %>%
  select(
    all_of(predictor_cols),
    Num_of_reviews
  )

set.seed(42) # For reproducibility

# Apply knnImputation to the focused DataFrame
imputed_data_num_reviews = knnImputation(temp_df_num_reviews, k = 5, meth = "median")

# Update Car_df_merged$Num_of_reviews
Car_df_merged$Num_of_reviews = imputed_data_num_reviews$Num_of_reviews

# Verify Num_of_reviews imputation
na_count_num_of_reviews = Car_df_merged$Num_of_reviews %>% is.na() %>% sum()
sprintf("Number of NA values in Num_of_reviews after KNN imputation: %d", na_count_num_of_reviews)

cat("Summary statistics for Num_of_reviews after imputation:\n")
Car_df_merged$Num_of_reviews %>% summary()
```

```{r}
predictor_cols = c("money", "Mileage", "Year", "Engine_Size", "Drivetrain",
                   "Fuel_type", "Transmission", "Cylinder_Configuration",
                   "Has_Turbo", "Has_GDI", "Has_DOHC", "Has_Supercharged",
                   "Num_of_reviews") # Include Num_of_reviews as a predictor

cat("\n--- Imputing General_rate ---\n")

# Create a focused temporary DataFrame for General_rate imputation
temp_df_general_rate = Car_df_merged %>%
  mutate(
    Drivetrain = as.factor(Drivetrain),
    Fuel_type = as.factor(Fuel_type),
    Transmission = as.factor(Transmission),
    Cylinder_Configuration = as.factor(Cylinder_Configuration)
  ) %>%
  select(
    all_of(predictor_cols),
    General_rate
  )

set.seed(42) # For reproducibility

# Apply knnImputation to the focused DataFrame
imputed_data_general_rate = knnImputation(temp_df_general_rate, k = 5, meth = "median")

# Update Car_df_merged$General_rate
Car_df_merged$General_rate = imputed_data_general_rate$General_rate

# Verify General_rate imputation
na_count_general_rate = Car_df_merged$General_rate %>%
  is.na() %>%
  sum()
sprintf("Number of NA values in General_rate after KNN imputation: %d", na_count_general_rate)

cat("Summary statistics for General_rate after imputation:\n")
Car_df_merged$General_rate %>% summary()
```

```{r}
write_csv(Car_df_merged, "car_df_merged.csv")
```

Due the massive amount of resources and time needed to use KNN imputation we decided to impute the rest with join multiple regression imputation

```{r}
Car_df_merged = read_csv("https://raw.githubusercontent.com/NitoBoritto/R_New_York_Car_Project/main/car_df_merged.csv")
```

```{r}
rating_cols = c(
  "Comfort",
  "Interior design",
  "Performance",
  "Value for the money",
  "Exterior styling",
  "Reliability"
)

predictor_cols = c("money", "Mileage", "Year", "Engine_Size",
                   "Drivetrain", "Fuel_type", "Transmission",
                   "Cylinder_Configuration","Num_of_reviews", "General_rate")

imp_vars <- c(predictor_cols, rating_cols)

imp_df <- Car_df_merged[, imp_vars] # Subset data

imp_df <- imp_df %>% # Convert categoricals
  mutate(across(
    c(Drivetrain, Fuel_type, Transmission, Cylinder_Configuration),
    as.factor
  ))

meth <- make.method(imp_df)
meth[rating_cols] <- "pmm"   # ratings get imputed using pmm (predective mean matching)
# all others left as default by mice

# Predictor matrix
pred <- make.predictorMatrix(imp_df)
diag(pred) <- 0              # so variables don't predict themselves


set.seed(42)

imp <- mice(
  imp_df,
  method = meth,
  predictorMatrix = pred,
  m = 5,
  maxit = 5,
  printFlag = TRUE
)

# ---- Use one completed dataset (ML / prediction) ----
Car_df_merged[, rating_cols] <- complete(imp, 1)[, rating_cols]


for (col in rating_cols) {
  na_count_col = Car_df_merged %>% pull(!!sym(col)) %>% is.na() %>% sum()

  cat(sprintf("Number of NA values in column after joint multiple imputation: %d\n", na_count_col))

  print("Summary statistics for column after imputation:\n")
  summary_stats = Car_df_merged %>% pull(!!sym(col)) %>% summary()
  print(summary_stats)
}

total_nulls = Car_df_merged %>% is.na() %>% sum()

if (total_nulls == 0) {
  print("Dataframe fully cleaned!")
}
```

```{r}
write_csv(Car_df_merged, "car_df_merged.csv")
```

### Outlier detection & handling

```{r}
numerical_cols_for_outliers = c('money', 'Mileage', 'Year', 'Num_of_reviews', 'General_rate', 'Comfort', 'Interior design', 'Performance', 'Value for the money', 'Exterior styling', 'Reliability', 'Engine_Size')

print("Numerical columns identified for outlier analysis:")
print(numerical_cols_for_outliers)
```

```{r}
outlier_bounds = list()

for (col_name in numerical_cols_for_outliers) {
  Q1 = Car_df_merged %>% pull(!!sym(col_name)) %>% quantile(0.25, na.rm = TRUE)
  Q3 = Car_df_merged %>% pull(!!sym(col_name)) %>% quantile(0.75, na.rm = TRUE)
  IQR = Q3 - Q1
  lower_bound = Q1 - 1.5 * IQR
  upper_bound = Q3 + 1.5 * IQR

  outlier_bounds = append(outlier_bounds, list(
    list(
      "column" = col_name,
      "Q1" = Q1,
      "Q3" = Q3,
      "IQR" = IQR,
      "lower_bound" = lower_bound,
      "upper_bound" = upper_bound
    )
  ))
}

print("Calculated outlier bounds for each numerical column:")
for (bound_info in outlier_bounds) {
  cat(sprintf(
    "Column: %s, Q1: %.2f, Q3: %.2f, IQR: %.2f, Lower Bound: %.2f, Upper Bound: %.2f\n",
    bound_info$column,
    bound_info$Q1,
    bound_info$Q3,
    bound_info$IQR,
    bound_info$lower_bound,
    bound_info$upper_bound
  ))
}
```

#### Winsorization

```{r}
for (bound_info in outlier_bounds) {
  col_name = bound_info$column
  lower_bound = bound_info$lower_bound
  upper_bound = bound_info$upper_bound

  # Apply capping and flooring
  Car_df_merged = Car_df_merged %>%
    mutate(!!sym(col_name) := pmax(pmin(!!sym(col_name), upper_bound), lower_bound))
}

cat("\nSummary statistics after outlier handling (capping and flooring):\n")
key_cols_to_summarize = c("money", "Mileage", "Engine_Size", "Year")
for (col in key_cols_to_summarize) {
  cat(sprintf("\n--- Column: %s ---\n", col))
  Car_df_merged %>% pull(!!sym(col)) %>% summary() %>% print()
}
```

```{r}
write_csv(Car_df_merged, "car_df_merged.csv")
```

## Dataset after cleaning

ONLY RUN THIS, NOTHING ELSE

```{r}
Car_df_merged = read_csv("https://raw.githubusercontent.com/NitoBoritto/R_New_York_Car_Project/main/car_df_merged.csv")
```

# EDA

use `+ theme_nito` after every `ggplot()` to style it up

```{r}
head(Car_df_merged)
```

```{r}
str(Car_df_merged)
```

```{r}
dim(Car_df_merged)
```

```{r}
colnames(Car_df_merged)
```

```{r}
summary(Car_df_merged["money"])
```

## Uni-Variate Analysis

### Price

```{r}
ggplot(Car_df_merged, aes(x = money)) +
  geom_histogram(bins = 30, fill = "steelblue", color = "black") +
  labs(title = "Distribution of Car Prices",
       x = "Price",
       y = "Count")
```

```{r}
ggplot(Car_df_merged, aes(y = money)) +
  geom_boxplot(fill = "tomato") +
  labs(title = "Price Outliers")
```

### Mileage

```{r}
theme_nito = theme(plot.title = element_text(hjust = 0.5, color = "darkgreen"),
    text = element_text(family = "garamond", size = 14),
	  rect = element_blank(),
	  panel.grid = element_blank(),
	  axis.line = element_line(color = "black"))


ggplot(Car_df_merged, aes(x = Mileage)) +
  geom_histogram(bins = 30, fill = "darkseagreen", color = "black") +
  labs(title = "Mileage Distribution") +
  theme_nito
```

### New vs Used

```{r}
ggplot(Car_df_merged, aes(x = `new&used`, fill = "slateblue")) +
  geom_bar() +
  labs(title = "New vs Used Cars",
       x = "new & used",
       y = "Count")
```

```{r}
new_used_data <- Car_df_merged %>%
  count(`new&used`) %>%
  mutate(percentage = n / sum(n))

ggplot(new_used_data, aes(x = "", y = percentage, fill = `new&used`)) +
  geom_bar(width = 1, stat = "identity") +
  coord_polar("y", start = 0) +
  geom_text(aes(label = scales::percent(percentage)), position = position_stack(vjust = 0.5)) +
  labs(title = "Distribution of New vs. Used Cars", fill = "Condition") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5, color = "blue", size = 16, face = "bold"),
        legend.position = "right") + theme_nito
```

### Fuel Type

```{r}
ggplot(Car_df_merged, aes(x = Fuel_type)) +
  geom_bar(fill = "orange") +
  labs(title = "Fuel Type Distribution") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Bivariate Analysis

### Mileage vs Price

```{r}
colnames(Car_df_merged)
```

```{r}
ggplot(Car_df_merged, aes(x = Mileage, y = money)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(title = "Mileage vs Price",
       x = "Mileage",
       y = "Price"
       )
```

### Condition vs Price

```{r}
ggplot(Car_df_merged, aes(x = `new&used`, y = money)) +
  geom_boxplot(fill = "lightblue") +
  labs(title = "Price by Condition")
```

### Fuel Type vs Price

```{r}
ggplot(Car_df_merged, aes(x = Fuel_type, y = money)) +
  geom_boxplot(fill = "khaki") +
  labs(title = "Fuel Type vs Price") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### Transmission vs Price

```{r}
ggplot(Car_df_merged, aes(x = Transmission, y = money)) +
  geom_boxplot(fill = "lightpink") +
  labs(title = "Transmission vs Price")
```

### Drivetrain Types vs Price

```{r}
ggplot(Car_df_merged, aes(x = Drivetrain, y = money, fill = Drivetrain)) +
  geom_boxplot() +
  scale_fill_brewer(palette = "Greens") +
  theme_minimal()+
  labs(title = "Distribution of Drivetrain Types")
```

### Brand vs Price

```{r}
ggplot(Car_df_merged, aes(x = reorder(brand, money, median), y = money)) +
  geom_boxplot() +
  coord_flip() +
  labs(title = "Price Distribution by Brand",
       x = "Brand",
       y = "Price")
```

## Important Binary Features

### Accidents or Damage

```{r}
ggplot(Car_df_merged, aes(x = factor(`Accidents or damage`), y = money)) +
  geom_boxplot(fill = "salmon", outlier.alpha = 0.3) +
  theme_minimal() +
  labs(
    title = "Impact of Accidents on Price",
    x = "Accident History",
    y = "Price"
  )
```

### Clean Title

```{r}
ggplot(Car_df_merged, aes(x = factor(`Clean title`), y = money)) +
  geom_boxplot(fill = "lightgreen", outlier.alpha = 0.3) +
  theme_minimal() +
  labs(
    title = "Clean Title vs Price",
    x = "Title Status",
    y = "Price"
  )
```

### One Owner

```{r}
ggplot(Car_df_merged, aes(x = factor(`1-owner vehicle`), y = money)) +
  geom_boxplot(fill = "lightgray", outlier.alpha = 0.3) +
  theme_minimal() +
  labs(
    title = "One Owner vs Price",
    x = "Ownership Status",
    y = "Price"
  )
```

### Year vs Price

```{r}
ggplot(Car_df_merged, aes(x = Year, y = money)) +
  geom_point(alpha = 0.5) +
  geom_smooth(se = FALSE, color = "blue") +
  labs(title = "Year vs Price")
```

## After Using Log

### Log Price

```{r}
ggplot(Car_df_merged, aes(x = log(money))) +
  geom_histogram(bins = 30, fill = "purple") +
  labs(title = "Log-Transformed Price Distribution")
```

### Log Mileage

```{r}
ggplot(Car_df_merged, aes(x = log(Mileage))) +
  geom_histogram(bins = 30, fill = "darkseagreen", color = "black") +
  labs(title = "Mileage Distribution")
```

### Log Mileage Vs Price

```{r}
ggplot(Car_df_merged, aes(x = log(Mileage), y = log(money))) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "lm") +
  labs(title = "Log Mileage vs Log Price")
```

### new.used vs log Price

```{r}
ggplot(Car_df_merged, aes(x = `new&used`, y = log(money))) +
  geom_boxplot(fill = "lightblue") +
  labs(title = "Log Price by Condition")
```

### Fuel Type Vs Log Price

```{r}
ggplot(Car_df_merged, aes(x = Fuel_type, y = log(money))) +
  geom_boxplot(fill = "khaki") +
  labs(title = "Fuel Type vs Price") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### Transmission Vs Log Price

```{r}
ggplot(Car_df_merged, aes(x = Transmission, y = log(money))) +
  geom_boxplot(fill = "lightpink") +
  labs(title = "Transmission vs Log Price")
```

### Drivertrain Types Vs Log Price

```{r}
ggplot(Car_df_merged, aes(x = Drivetrain, y = log(money), fill = Drivetrain)) +
  geom_boxplot() +
  scale_fill_brewer(palette = "Greens") +
  theme_minimal()+
  labs(title = "Distribution of Drivetrain Types")
```

### Brand Vs Log Price

```{r}
ggplot(Car_df_merged, aes(x = reorder(brand, money, median), y = log(money))) +
  geom_boxplot() +
  coord_flip() +
  labs(title = "Price Distribution by Brand",
       x = "Brand",
       y = "Price")
```

### Accidents or Damage

```{r}
ggplot(Car_df_merged, aes(x = factor(`Accidents or damage`), y = log(money))) +
  geom_boxplot(fill = "salmon", outlier.alpha = 0.3) +
  theme_minimal() +
  labs(
    title = "Impact of Accidents on Price",
    x = "Accident History",
    y = "Price"
  )
```

### Clean Title

```{r}
ggplot(Car_df_merged, aes(x = factor(`Clean title`), y = log(money))) +
  geom_boxplot(fill = "lightgreen", outlier.alpha = 0.3) +
  theme_minimal() +
  labs(
    title = "Clean Title vs Price",
    x = "Title Status",
    y = "Price"
  )
```

### One Owner

```{r}
ggplot(Car_df_merged, aes(x = factor(`1-owner vehicle`), y = log(money))) +
  geom_boxplot(fill = "lightgray", outlier.alpha = 0.3) +
  theme_minimal() +
  labs(
    title = "One Owner vs Price",
    x = "Ownership Status",
    y = "Price"
  )
```

### Year Vs Price

```{r}
ggplot(Car_df_merged, aes(x = Year, y = log(money))) +
  geom_point(alpha = 0.5) +
  geom_smooth(se = FALSE, color = "blue") +
  labs(title = "Year vs Price")
```

## Correltion Matrix

### Correlation Matrix between Year, Price, and Mileage

```{r}
corr_matrix <- cor(Car_df_merged[, c("Year", "money", "Mileage")], use = "complete.obs")

melted_corr <- melt(corr_matrix)

ggplot(data = melted_corr, aes(x=Var1, y=Var2, fill=value)) +
  geom_tile() +
  geom_text(aes(label = round(value, 2)), color = "white") +
  scale_fill_gradient2(low = "#063970", high = "#f5d142", mid = "#7a7d7d",
                       midpoint = 0, limit = c(-1,1)) +
  theme_minimal() +
  labs(title = "Distribution of Vehicle Drivetrains",x = "", y = "")
```

### Correlation Matrix all Data

```{r}
target_cols <- c("money", "Mileage", "Year", "MPG", "Engine_Size",
                 "Num_of_reviews", "General_rate", "Comfort",
                 "Interior design", "Performance", "Value for the money",
                 "Exterior styling", "Reliability")

full_corr <- cor(Car_df_merged[, target_cols], use = "complete.obs")


ggcorrplot(full_corr,
           hc.order = TRUE,
           type = "lower",
           lab = TRUE,
           lab_size = 2.5,
           method = "square",
           colors = c("#d73027", "white", "#1a9850"), # Red (negative) to Green (positive)
           title = "Full Vehicle Data Correlation Analysis",
           outline.col = "white") +
  theme(
    axis.text.x = element_text(size = 8, angle = 45, vjust = 1, hjust = 1),
    axis.text.y = element_text(size = 8),
    plot.title = element_text(size = 12, face = "bold", hjust = 0.5)
  )
```

## Relationship

### Relationship between Manufacturing Year, Price, and Mileage

```{r}
cols_to_plot <- c("Year", "money", "Mileage")

ggpairs(Car_df_merged[, cols_to_plot]) +
  theme_bw() +
  labs(
    title = "Vehicle Attribute Pair Plot",
    subtitle = "Relationship between Manufacturing Year, Price, and Mileage"
  ) +
  theme(plot.title = element_text(face = "bold", size = 14))
```

# Regression

## Simple Linear Regression

```{r}
money_Mileage_model <- lm(money ~ Mileage, data = Car_df_merged)
```

```{r}
summary(money_Mileage_model)
```

```{r}
autoplot(money_Mileage_model,which=1:3,nrow=3,ncol=1)
```

```{r}
money_Mileage_model_logged <- lm(log(money) ~ Mileage, data = Car_df_merged)
summary(money_Mileage_model_logged)
```

## Multiple Regression

### Dummy Variables

```{r}
Car_df_merged <- dummy_cols(Car_df_merged, select_columns = c("brand", "Transmission", "Fuel_type"),
remove_first_dummy = TRUE,
remove_selected_columns = TRUE)
```

```{r}
# Create the tibble by removing the target variable (money)
explanatory_data <- Car_df_merged %>%
  select(-money)
print(explanatory_data)
```

```{r}
explanatory_data_numeric <- Car_df_merged %>%
  select(where(is.numeric)) %>%  # Keeps only numeric columns
  select(-money)                 # Removes the target variable

# Check which columns were included
colnames(explanatory_data_numeric)
```

```{r}
ggplot(data = Car_df_merged, aes(x = money)) +
  geom_histogram()
```

### Scaling the Data (using Standardization Method)

```{r}
# 1. Prepare the data by selecting relevant columns
# We use drop_na() because most models will fail if there are missing values
model_ready_data <- Car_df_merged %>%
  select(money, Mileage, Year, Engine_Size, MPG,
         starts_with("brand_"), starts_with("Transmission_"),`Accidents or damage`) %>%
  drop_na()

# 2. Scale only the continuous predictors
# This centers (mean=0) and scales (SD=1) the numbers
Car_df_final <- model_ready_data %>%
  mutate(across(c(Mileage, Year, Engine_Size, MPG), ~ as.vector(scale(.))))

# View the result
head(Car_df_final)
```

### Check the Scaling

```{r}
# This will show the mean and SD for your numeric columns
Car_df_final %>%
  summarise(across(c(Mileage, Year, Engine_Size, MPG),
                   list(mean = ~round(mean(.), 2),
                        sd = ~round(sd(.), 2))))
```

# Model

## Model –\> (1)

```{r}
# Build a model with all potential predictors
# 'money' is the target, all others are explanatory
model_1 <- lm(money ~ Mileage  +Year + Engine_Size + MPG +  brand_Toyota + brand_BMW + Transmission_Variable+`Accidents or damage`, data = Car_df_final)

# View the full statistical report
summary(model_1)
```

#### Check Multi Correinality

```{r}
# Calculate Variance Inflation Factor (VIF)
vif(model_1)
```

```{r}
autoplot(model_1,which=1:3,nrow=3,ncol=1)
```

## Model –\> (2) (Transformation + Interaction)

```{r}
# 1. Transform Raw Data
model_ready <- Car_df_merged %>%
  # Ensure these columns exist
  drop_na(money, Mileage, Year, Engine_Size, MPG) %>%
  mutate(
    log_money = log(money),
    log_mileage = log(Mileage + 1),
    log_engine = log(Engine_Size + 0.1)
  )

# 2. Scale the Continuous Features
final_data <- model_ready %>%
  # FIXED: Removed Age and Age_Sq from the scaling list
  mutate(across(c(log_mileage, log_engine, MPG),
                ~ as.vector(scale(.)))) %>%
  # FIXED: Removed the extra closing parenthesis
  rename_with(~make.names(.))

# 3. Fit Model with Interactions

model_final <- lm(log_money ~
                    log_mileage * brand_BMW +
                    brand_Toyota +
                    log_engine +
                    Transmission_Variable +
                    MPG,
                  data = final_data)

# 4. View results
summary(model_final)
```

## Model –\> (3) (Feature Engineering)

```{r}
# 1. Transform Raw Data
model_ready <- Car_df_merged %>%

  drop_na(money, Mileage, Year, Engine_Size, MPG) %>%
  mutate(
    log_money = log(money),
    log_mileage = log(Mileage + 1),
    # Use + 0.1 so that Electric cars (0.0 engine) don't become -Inf
    log_engine = log(Engine_Size + 0.1),
    Age = 2025 - Year, Age_Sq = Age^2 )

# 2. Scale the Continuous Features
final_data <- model_ready %>%
  mutate(across(c(log_mileage, log_engine, Age, Age_Sq, MPG),
                ~ as.vector(scale(.)))) %>%
  rename_with(~make.names(.))

# 3. Fit Model with Interactions

model_final <- lm(log_money ~ log_mileage * brand_BMW + brand_Toyota + log_engine +Transmission_Variable+ Age_Sq + MPG, data = final_data)


# 4. View results
summary(model_final)
```

```{r}
# Calculate Variance Inflation Factor (VIF)
vif(model_final)
```

```{r}
autoplot(model_1,which=1:3,nrow=3,ncol=1)
```

## Model –\> (4) (Random Forest)

```{r}
# 1. Prepare Data and Create Age
rf_data <- Car_df_merged %>%
  distinct() %>% # Prevent record leakage as no car in be in both trianing and test data
  mutate(
    Age = 2025 - Year # Create Age from Year
  ) %>%
  select(money, Mileage, Age, Engine_Size, MPG,
         starts_with("brand_"), starts_with("Transmission_")) %>%
  drop_na()

# Clean column names
names(rf_data) <- make.names(names(rf_data))

# 2. Split Data BEFORE scaling
set.seed(123)
trainIndex <- createDataPartition(rf_data$money, p = .8, list = FALSE)
train_raw <- rf_data[trainIndex, ]
test_raw  <- rf_data[-trainIndex, ]

# 3. Scale Continuous Variables (using Train parameters for Test)

cont_cols <- c("Mileage", "Age", "Engine_Size", "MPG")

# Train Scaling
train_data <- train_raw %>%
  mutate(
    money = log(money), # Log target for better RF performance
    across(all_of(cont_cols), ~ as.vector(scale(.)))
  )

# Extract scaling parameters to prevent leakage
means <- colMeans(train_raw[, cont_cols])
sds   <- apply(train_raw[, cont_cols], 2, sd)

# Test Scaling (Applying train parameters)
test_data <- test_raw %>%
  mutate(
    money = log(money),
    Mileage = (Mileage - means["Mileage"]) / sds["Mileage"],
    Age = (Age - means["Age"]) / sds["Age"],
    Engine_Size = (Engine_Size - means["Engine_Size"]) / sds["Engine_Size"],
    MPG = (MPG - means["MPG"]) / sds["MPG"]
  )

# 4. Build Model
rf_model <- randomForest(money ~ ., data = train_data, ntree = 100, importance = TRUE)
print(rf_model)
```

# Review model

```{r}
varImpPlot(rf_model, main = "Variable Importance: Is Age a Top Predictor?")
```

## RMSE & Model Evaluation

```{r}
# 1. Predict using the Test set (results will be in log scale)
predictions_log <- predict(rf_model, test_data)

# 2. Transform values from Log scale back to actual Dollars using exp()
predictions_dollars <- exp(predictions_log)
actual_dollars <- exp(test_data$money)

# 3. Calculate Root Mean Squared Error (RMSE) in dollars
rmse_dollars <- sqrt(mean((actual_dollars - predictions_dollars)^2))

# 4. Calculate Mean Absolute Error (MAE) in dollars
# This tells you: On average, how many dollars is the model off by?
mae_dollars <- mean(abs(actual_dollars - predictions_dollars))

# Print the real-world error metrics
cat("--- Real Error Metrics (In Dollars) ---\n")
cat("Average Error (MAE): $", round(mae_dollars, 2), "\n")
cat("Root Mean Squared Error (RMSE): $", round(rmse_dollars, 2), "\n")

# 5. Create a plot to compare Actual vs. Predicted prices
plot_data <- data.frame(Actual = actual_dollars, Predicted = predictions_dollars)

ggplot(plot_data, aes(x = Actual, y = Predicted)) +
  geom_point(alpha = 0.3, color = "blue") +
  # Add a 45-degree dashed red line representing perfect predictions
  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
  scale_x_continuous(labels = label_dollar()) +
  scale_y_continuous(labels = label_dollar()) +
  labs(
    title = "Actual vs. Predicted Car Prices",
    subtitle = paste("Mean Absolute Error: $", round(mae_dollars, 2)),
    x = "Actual Price ($)",
    y = "Predicted Price ($)"
  ) +
  theme_minimal()
```

# Simulation

## Checking structure

```{r}
str(Car_df_merged)
```

## Numerical simulation

```{r}
set.seed(42)

Car_df_sampled = Car_df_merged %>%
  sample_n(50000)

print("Dimensions of the sampled DataFrame (Car_df_sampled):")
Car_df_sampled %>% dim()

print("First few rows of Car_df_sampled:")
Car_df_sampled %>% head()
```

```{r}
set.seed(42)

Sim_Newmerc_Sampled_list = list()
num_samples_sampled = nrow(Car_df_sampled) # Use the number of rows from the sampled data

# Helper function to ensure values are within realistic/original bounds after inverse transformation
clamp = function(x, min_val, max_val) {
  pmax(pmin(x, max_val), min_val)
}

# 1. money: Log transformation
original_money_sampled = Car_df_sampled$money
transformed_money_sampled = log(original_money_sampled)
mean_money_sampled = mean(transformed_money_sampled)
sd_money_sampled = sd(transformed_money_sampled)
sim_transformed_money_sampled = rnorm(num_samples_sampled, mean = mean_money_sampled, sd = sd_money_sampled)
sim_money_sampled = exp(sim_transformed_money_sampled)
Sim_Newmerc_Sampled_list$money = clamp(sim_money_sampled, min(original_money_sampled), max(original_money_sampled))

# 2. Mileage: Log(x+1) transformation
original_mileage_sampled = Car_df_sampled$Mileage
transformed_mileage_sampled = log(original_mileage_sampled + 1)
mean_mileage_sampled = mean(transformed_mileage_sampled)
sd_mileage_sampled = sd(transformed_mileage_sampled)
sim_transformed_mileage_sampled = rnorm(num_samples_sampled, mean = mean_mileage_sampled, sd = sd_mileage_sampled)
sim_mileage_sampled = exp(sim_transformed_mileage_sampled) - 1
Sim_Newmerc_Sampled_list$Mileage = clamp(sim_mileage_sampled, min(original_mileage_sampled), max(original_mileage_sampled))

# 3. Year: No transformation
original_year_sampled = Car_df_sampled$Year
mean_year_sampled = mean(original_year_sampled)
sd_year_sampled = sd(original_year_sampled)
sim_year_sampled = rnorm(num_samples_sampled, mean = mean_year_sampled, sd = sd_year_sampled)
sim_year_sampled = round(clamp(sim_year_sampled, min(original_year_sampled), max(original_year_sampled)))
Sim_Newmerc_Sampled_list$Year = sim_year_sampled

# 4. MPG: Log(x+1) transformation
original_mpg_sampled = Car_df_sampled$MPG
transformed_mpg_sampled = log(original_mpg_sampled + 1)
mean_mpg_sampled = mean(transformed_mpg_sampled)
sd_mpg_sampled = sd(transformed_mpg_sampled)
sim_transformed_mpg_sampled = rnorm(num_samples_sampled, mean = mean_mpg_sampled, sd = sd_mpg_sampled)
sim_mpg_sampled = exp(sim_transformed_mpg_sampled) - 1
Sim_Newmerc_Sampled_list$MPG = clamp(sim_mpg_sampled, min(original_mpg_sampled), max(original_mpg_sampled))

# 5. Num_of_reviews: Log(x+1) transformation
original_num_reviews_sampled = Car_df_sampled$Num_of_reviews
transformed_num_reviews_sampled = log(original_num_reviews_sampled + 1)
mean_num_reviews_sampled = mean(transformed_num_reviews_sampled)
sd_num_reviews_sampled = sd(transformed_num_reviews_sampled)
sim_transformed_num_reviews_sampled = rnorm(num_samples_sampled, mean = mean_num_reviews_sampled, sd = sd_num_reviews_sampled)
sim_num_reviews_sampled = exp(sim_transformed_num_reviews_sampled) - 1
sim_num_reviews_sampled = round(clamp(sim_num_reviews_sampled, min(original_num_reviews_sampled), max(original_num_reviews_sampled)))
Sim_Newmerc_Sampled_list$Num_of_reviews = sim_num_reviews_sampled

# 6. Rating columns: No transformation
rating_cols_no_transform = c('General_rate', 'Comfort', 'Interior design', 'Performance', 'Value for the money', 'Exterior styling', 'Reliability')
for (col in rating_cols_no_transform) {
  original_data_sampled = Car_df_sampled[[col]]
  mean_data_sampled = mean(original_data_sampled)
  sd_data_sampled = sd(original_data_sampled)
  sim_data_sampled = rnorm(num_samples_sampled, mean = mean_data_sampled, sd = sd_data_sampled)
  Sim_Newmerc_Sampled_list[[col]] = round(clamp(sim_data_sampled, min(original_data_sampled), max(original_data_sampled)), 1)
}

# 7. Engine_Size: Log(x+0.1) transformation
original_engine_size_sampled = Car_df_sampled$Engine_Size
transformed_engine_size_sampled = log(original_engine_size_sampled + 0.1)
mean_engine_size_sampled = mean(transformed_engine_size_sampled)
sd_engine_size_sampled = sd(transformed_engine_size_sampled)
sim_transformed_engine_size_sampled = rnorm(num_samples_sampled, mean = mean_engine_size_sampled, sd = sd_engine_size_sampled)
sim_engine_size_sampled = exp(sim_transformed_engine_size_sampled) - 0.1
Sim_Newmerc_Sampled_list$Engine_Size = clamp(sim_engine_size_sampled, min(original_engine_size_sampled), max(original_engine_size_sampled))

# Combine into a data frame
Sim_Newmerc_Sampled = as.data.frame(Sim_Newmerc_Sampled_list)

cat("Sim_Newmerc_Sampled DataFrame created successfully. First few rows:\n")
print(head(Sim_Newmerc_Sampled))

cat("Summary statistics for Sim_Newmerc_Sampled:\n")
print(summary(Sim_Newmerc_Sampled))
```

```{r}
if (!exists("theme_nito")) {
  theme_nito = theme(plot.title = element_text(hjust = 0.5, color = "blue"),
      text = element_text(family = "sans", size = 14),
      rect = element_blank(),
      panel.grid = element_blank(),
      axis.line = element_line(color = "black"))
}

key_cols_for_comparison_sampled = c('money', 'Mileage', 'MPG', 'Engine_Size', 'Num_of_reviews')

plot_list_sampled = list()

for (col_name in key_cols_for_comparison_sampled) {
  # Create a combined dataframe for plotting
  df_combined_sampled = data.frame(
    value = c(Car_df_sampled[[col_name]], Sim_Newmerc_Sampled[[col_name]]),
    source = factor(c(rep("Original Sampled", nrow(Car_df_sampled)), rep("Simulated Sampled", nrow(Sim_Newmerc_Sampled))))
  )

  # Generate density plot
  p_sampled = ggplot(df_combined_sampled, aes(x = value, fill = source, color = source)) +
    geom_density(alpha = 0.5) +
    labs(
      title = paste("Distribution of", col_name, "(Original Sampled vs. Simulated Sampled)"),
      x = col_name,
      y = "Density",
      fill = "Data Source",
      color = "Data Source"
    ) +
    theme_nito +
    theme(legend.position = "top")

  plot_list_sampled[[col_name]] = p_sampled
}

# Arrange plots in a grid
grid.arrange(grobs = plot_list_sampled, ncol = 2,
             top = textGrob("Comparison of Original Sampled vs. Simulated Sampled Numerical Distributions",
                            gp = gpar(fontsize = 16, col = "darkgreen", fontface = "bold")))
```

## Categorical simulation

```{r}
categorical_cols_for_simulation_sampled = c('new&used', 'Drivetrain', 'brand', 'Fuel_type', 'Transmission', 'Cylinder_Configuration')

Simulated_Categorical_Data_Sampled = data.frame(row.names = 1:nrow(Car_df_sampled))

for (col_name in categorical_cols_for_simulation_sampled) {
  cat(sprintf("\n--- Analysis and Simulation for: %s (Sampled Data) ---\n", col_name))

  # 1. Analyze Original Sampled Distribution (Frequency Table)
  cat("Original Sampled Frequency Table:\n")
  original_table_sampled = table(Car_df_sampled[[col_name]])
  print(original_table_sampled)
  cat("\nOriginal Sampled Proportion Table:\n")
  original_prop_table_sampled = prop.table(original_table_sampled)
  print(original_prop_table_sampled)

  # 2. Simulate New Data based on proportions from sampled data
  set.seed(42) # For reproducibility of simulation
  simulated_values_sampled = sample(
    x = names(original_prop_table_sampled),
    size = nrow(Car_df_sampled),
    replace = TRUE,
    prob = original_prop_table_sampled
  )
  Simulated_Categorical_Data_Sampled[[col_name]] = simulated_values_sampled

  # 3. Analyze Simulated Distribution (Frequency Table)
  cat("\nSimulated Sampled Frequency Table:\n")
  simulated_table_sampled = table(Simulated_Categorical_Data_Sampled[[col_name]])
  print(simulated_table_sampled)
  cat("\nSimulated Sampled Proportion Table:\n")
  simulated_prop_table_sampled = prop.table(simulated_table_sampled)
  print(simulated_prop_table_sampled)
  cat("\n") # Add a newline for better separation between columns
}

cat("\n--- First few rows of the complete Simulated_Categorical_Data_Sampled DataFrame ---\n")
print(head(Simulated_Categorical_Data_Sampled))

cat("\n--- Structure of the complete Simulated_Categorical_Data_Sampled DataFrame ---\n")
str(Simulated_Categorical_Data_Sampled)
```

```{r}
if (!exists("theme_nito")) {
  theme_nito = theme(plot.title = element_text(hjust = 0.5, color = "blue"),
      text = element_text(family = "sans", size = 14),
      rect = element_blank(),
      panel.grid = element_blank(),
      axis.line = element_line(color = "black"))
}

categorical_cols_for_comparison_sampled = c('new&used', 'Drivetrain', 'brand', 'Fuel_type', 'Transmission', 'Cylinder_Configuration')

plot_list_sampled_categorical = list()

for (col_name in categorical_cols_for_comparison_sampled) {
  # Create combined dataframe for plotting
  df_combined_sampled_cat = bind_rows(
    Car_df_sampled %>% select(!!sym(col_name)) %>% mutate(source = "Original Sampled"),
    Simulated_Categorical_Data_Sampled %>% select(!!sym(col_name)) %>% mutate(source = "Simulated Sampled")
  ) %>%
  # Ensure the category column is a factor for consistent plotting
  mutate(!!sym(col_name) := as.factor(!!sym(col_name)))

  # Calculate proportions
  df_proportions_sampled_cat = df_combined_sampled_cat %>%
    group_by(source, !!sym(col_name)) %>%
    summarise(count = n(), .groups = 'drop') %>%
    group_by(source) %>%
    mutate(proportion = count / sum(count))

  # Generate bar plot
  p_sampled_cat = ggplot(df_proportions_sampled_cat, aes(x = !!sym(col_name), y = proportion, fill = source)) +
    geom_bar(stat = "identity", position = "dodge") +
    labs(
      title = paste("Distribution of", col_name),
      x = col_name,
      y = "Proportion",
      fill = "Data Source"
    ) +
    theme_nito +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

  plot_list_sampled_categorical[[col_name]] = p_sampled_cat
}

# Arrange plots in a grid
num_plots_cat = length(plot_list_sampled_categorical)
num_cols_cat = min(3, ceiling(num_plots_cat / 2)) # Max 3 columns

grid.arrange(grobs = plot_list_sampled_categorical, ncol = num_cols_cat,
             top = textGrob("Comparison of Original Sampled vs. Simulated Sampled Categorical Distributions",
                            gp = gpar(fontsize = 16, col = "darkgreen", fontface = "bold")))
```

```{r}
Simulated_Data_Sampled = cbind(Sim_Newmerc_Sampled, Simulated_Categorical_Data_Sampled)

cat("\n--- First few rows of the complete Simulated_Data_Sampled DataFrame ---\n")
print(head(Simulated_Data_Sampled))

cat("\n--- Structure of the complete Simulated_Data_Sampled DataFrame ---\n")
str(Simulated_Data_Sampled)
```
